Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries

### 2021-10-27 Manuscript/manuscript.tex
Finished Discussion and submitted to preprint server

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

  \title{Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada}
    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
   \ead{samuel.robinson@ucalgary.ca} 
    \author[University of Calgary]{Lan H. Nguyen}
   \ead{hoanglan.nguyen@ucalgary.ca} 
    \author[University of Calgary]{Paul Galpern}
   \ead{paul.galpern@ucalgary.ca} 
      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~moisture, temperature), and can act as refuges for beneficial arthropods. This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.~pollination or pest suppression) from the field boundary. However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure. Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers. In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types. Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance. There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but this was not seen for other crop types or boundary types. This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
  \end{abstract}
   \begin{keyword} precision yield; spillover; landscape;\end{keyword}
 \end{frontmatter}

\newpage
\doublespacing

@@

SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this depends on both the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Kowalchuk \& Jong 1995; Garibaldi \emph{et al.} 2011).
Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
Yield may also be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services.
However, this scenario is rarely considered within studies of SNL or boundary effects (but see Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance. Scenario 1: no effect of boundary. Scenario 2: negative effect on yield. Scenario 3: positive effect on yield. Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
\end{figure}

Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.\\
Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
Meta-analyses are a common solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs.
However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
1. How does crop yield change with distance from the edge of field?
2. Does this depend on type of field edge?
3. Is there an intermediate distance where yield is maximized or variance is minimized?
To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.

@@

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
Therefore, we used the following filtering procedure to clean the data prior to model fitting:

@@

Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
6. Grassland: permanent seminatural grassland or pasture (not in rotation)

@@

\begin{equation}
  \begin{split}
  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60)\\
  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=60)\\
  \end{split}
  \end{equation}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Yield = weight of crop gathered per area of harvest (T/ha)
\item
  Boundary Distance = distance from field boundary type \emph{i} (m)
\item
  Easting, Northing = distance from centre of field (m)
\item
  Sequence = order of harvest within field (1--N points)
\item
  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
\end{enumerate}

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).

The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
That is, combining of fields occurs in patterns (e.g.~east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.

To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
2. Use posterior sample to simulate a new smoother from each field
3. Fit model of simulated smoothers from all fields, and save this ``meta-smoother''
4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field. Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.}\label{fig:spatialSmooths}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
\end{figure}

Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).

@@

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.

@@

@@

@@

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\AttributeTok}[1]{#1}
\newcommand{\BaseNTok}[1]{#1}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\ConstantTok}[1]{#1}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
\newcommand{\DataTypeTok}[1]{#1}
\newcommand{\DecValTok}[1]{#1}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{#1}
\newcommand{\FunctionTok}[1]{#1}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{#1}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
\newcommand{\VariableTok}[1]{#1}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc header
\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} %\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

  \title{Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries}
    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
   \ead{samuel.robinson@ucalgary.ca} 
    \author[University of Calgary]{Lan H. Nguyen}
   \ead{hoanglan.nguyen@ucalgary.ca} 
    \author[University of Calgary]{Paul Galpern}
   \ead{paul.galpern@ucalgary.ca} 
      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  Field boundaries can provide ecosystem services to crops by creating better abiotic conditions for crop growth, and can also act as habitat for beneficial arthropods. This suggests that crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services from the field boundary. However, there is little large-scale evidence showing this, largely because plot-scale crop yields are costly and time-consuming to measure. Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is widespread and data is frequently recorded by growers. In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types. Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance. There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but this was not seen for other crop types or boundary types. This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
  \end{abstract}
   \begin{keyword} precision yield; spillover; landscape; shelterbelt;\end{keyword}
 \end{frontmatter}

\newpage

\onehalfspacing

@@

SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this depends on both the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than larger agricultural set-asides.

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance. Scenario 1: no effect of boundary. Scenario 2: negative edge effect on yield. Scenario 3: positive ecosystem service effect on yield. Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
\end{figure}

Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Brandle \emph{et al.} 1984; Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
Meta-analyses are a common solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Osorio \emph{et al.} 2018).
However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
1. How does crop yield change with distance from the edge of field?
2. Does this depend on type of field edge?
3. Is there an intermediate distance where yield is maximized or variance is minimized?
To our knowledge, only (Osorio \emph{et al.} 2018) has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.

@@

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):

@@

Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
\item
  Wetland: permanent wetland; borders are largely unchanged from year-to-year
\item
  Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
\item
  Other crop: annual crop or pasture with little or no visible boundary between planted areas
\item
  Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
\item
  Grassland: permanent seminatural grassland or pasture (not in rotation)
\end{enumerate}

@@

\begin{equation}
  \begin{split}
  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=80)\\
  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
   & f(\text{Easting}, \text{Northing}, b=80)\\
  \end{split}
  \end{equation}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Yield = weight of crop gathered per area of harvest (T/ha)
\item
  Boundary Distance = distance from field boundary type \emph{i} (m)
\item
  Easting, Northing = distance from centre of field (m)
\item
  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
\end{enumerate}

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).

The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
That is, combining of fields occurs in patterns (e.g.~east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.

To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
\item
  Use posterior sample to simulate a new smoother from each field.
\item
  Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
\item
  Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
  This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
  Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
\end{enumerate}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field. Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field. Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}\label{fig:spatialSmooths}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields. The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}\label{fig:boundaryTypes}
\end{figure}

Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).

@@

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.

@@

\leavevmode\hypertarget{ref-bao2012}{}%
Bao, Y., Li, H. \& Zhao, H. (2012). Effect of shelterbelts on winter wheat yields in sanded farmland of north-western shandong province, china. \emph{Journal of Food Agriculture \& Environment}, 10, 1399--1403.

@@

\leavevmode\hypertarget{ref-brandle1984}{}%
Brandle, J.R., Johnson, B.B. \& Dearmont, D.D. (1984). Windbreak economics: The case of winter wheat production in eastern {Nebraska}. \emph{Journal of Soil and Water Conservation}, 39, 339--343.

@@

\leavevmode\hypertarget{ref-deJong1995}{}%
de Jong, E. \& Kowalchuk, T.E. (1995). The effect of shelterbelts on erosion and soil properties. \emph{Soil Science}, 159, 337--345.

\leavevmode\hypertarget{ref-dong2015}{}%
Dong, Z., Ouyang, F., Lu, F. \& Ge, F. (2015). Shelterbelts in agricultural landscapes enhance ladybeetle abundance in spillover from cropland to adjacent habitats. \emph{BioControl}, 60, 351--361.

@@

\singlespacing
\nolinenumbers

\newpage
